---
title: "10_learn_tmaptools"
author: "Author: [Steve, Yu](https://github.com/littlefish0331)"
date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%Y %b %d %a, %H:%M:%S')`" 
output:
  rmdformats::readthedown:
    css: style.css
    self_contained: TRUE
    thumbnails: FALSE
    lightbox: TRUE
    gallery: FALSE
    highlight: tango #探戈橘
    code_folding: show
    toc_depth: 3
    
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
rm(list = ls()); gc()
library(knitr)
library(kableExtra)
library(dplyr)
library(data.table)
library(tmap)
library(tmaptools)
knitr::opts_chunk$set(
	# 這邊是針對所有chunk的設定
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# File Target

- learn tmptoola package

裡面有很多函式是基於 [osm-search/Nominatim: Open Source search based on OpenStreetMap data](https://github.com/osm-search/Nominatim)

---

# Tool functions (shape)

## approx_areas

NLD_muni: World and Netherlands map，世界和荷蘭地圖

```{r}
data(NLD_muni)
NLD_muni %>% head
```

```{r}
NLD_muni$area <- approx_areas(NLD_muni, total.area = 33893)
NLD_muni %>% head
```

```{r}
tm_shape(NLD_muni) + 
  tm_bubbles(size="area", title.size = expression("Area in " * km^2))
```

```{r}
# function that returns min, max, mean and sum of area values
summary_areas <- function(x) {
  list(min_area = min(x),
       max_area = max(x),
       mean_area = mean(x),
       sum_area = sum(x))
}
```

```{r}
# area of the polygons
# approx_areas(NLD_muni) %>% head
# approx_areas(NLD_muni) %>% str
approx_areas(NLD_muni) %>% summary_areas()
```

```{r}
# area of the polygons, adjusted corrected for a specified total area size
# 多邊形的面積，調整為指定的總面積大小
approx_areas(NLD_muni, total.area = 33893) %>% summary_areas()
```


```{r}
# proportional area of the polygons
approx_areas(NLD_muni, target = "prop") %>% summary_areas()
```

```{r}
# area in squared miles
approx_areas(NLD_muni, target = "mi mi") %>% summary_areas()
```

```{r}
# sf:::st_proj_info()

# ---
# area of the polygons when unprojected
# NLD_muni %>% sf::st_transform(crs = 4326) %>% head
# NLD_muni %>% head
NLD_muni %>% sf::st_transform(crs = 4326) %>% approx_areas() %>% summary_areas()
```

## approx_distances

```{r}
data(NLD_prov)
```

```{r}
# North-South and East-West distances of the Netherlands
approx_distances(NLD_prov)
approx_distances(NLD_prov) %>% str
```

```{r}
# Distance between Maastricht and Groningen
p_maastricht <- geocode_OSM("Maastricht")$coords
p_groningen <- geocode_OSM("Groningen")$coords
approx_distances(p_maastricht, p_groningen, projection = 4326, target = "km")
# approx_distances(p_maastricht, p_groningen, target = "km") 
# 沒有指定projection，會變成計算普通直角坐標距離。
```

```{r}
# Check distances in several projections
# 可知投影真的有差
sapply(c(3035, 28992, 4326), function(projection) {
  p_maastricht <- geocode_OSM("Maastricht", projection = projection)$coords
  p_groningen <- geocode_OSM("Groningen", projection = projection)$coords
  approx_distances(p_maastricht, p_groningen, projection = projection)
})
```

## bb

看不太懂 參數relative 的作用。

```{r}
## load shapes
data(NLD_muni)
data(World)
```

```{r}
## get bounding box (similar to sp's function bbox)
bb(NLD_muni)
bb(NLD_muni) %>% str
```

```{r}
## extent it by factor 1.10
bb(NLD_muni, ext = 1.10)
```

```{r}
## convert to longlat
bb(NLD_muni, projection = 4326)
```

```{r}
## change existing bounding box
bb(NLD_muni, ext=1.5)
bb(NLD_muni, width=2, relative = TRUE)
bb(NLD_muni, xlim=c(.25, .75), ylim=c(.25, .75), relative = TRUE)
bb(NLD_muni, xlim=c(.25, .75), ylim=c(.25, .75), relative = F)
```

## bb_poly, bb_earth

```{r}
data(NLD_muni)
```

```{r}
current.mode <- tmap_mode("view")
qtm(bb_poly(NLD_muni))
```

```{r}
# restore mode
tmap_mode(current.mode)
```

## get_asp_ratio

```{r}
data(World)
get_asp_ratio(World)
```

```{r}
bb(World)
get_asp_ratio(bb(World))
```


```{r}
tm <- qtm(World)
get_asp_ratio(tm)
```

---

# Spatial transformation functions

## crop_shape

有底像是交集的動作。

```{r}
data(World, NLD_muni, land, metro)
qtm(land, raster = "trees", style = "natural")
```

```{r}
# 參數polygon = T，表示區塊的中心要重疊在邊界內才會留下來。
tmap_mode("view")
land_NLD <- crop_shape(land, NLD_muni, polygon = T)
qtm(land_NLD, raster = "trees", style = "natural")
```

```{r}
# 參數polygon = F，表示區塊只要有碰到邊界，就會算。
land_NLD <- crop_shape(land, NLD_muni, polygon = F)
qtm(land_NLD, raster = "trees", style = "natural")
```

```{r}
metro_Europe <- crop_shape(x = metro, 
                           y = World[World$continent == "Europe", ], 
                           polygon = TRUE)
qtm(World) +
  tm_shape(metro_Europe) +
  tm_bubbles("pop2010", col="red", title.size="European cities") +
  tm_legend(frame = TRUE)
# ttm()
```

## simplify_shape

```{r}
data(World)
World %>% head
```

```{r eval = F}
# show different simplification factors
# simplify_shape(): 函式沒有包好~ 目前失效。
jsonlite::toJSON(World, force = T) %>% str

# ---
tm1 <- qtm(World %>% simplify_shape(shp = ., fact = 0.05, force_FC = F), title = "Simplify 0.05")
tm2 <- qtm(World %>% simplify_shape(fact = 0.1), title = "Simplify 0.1")
tm3 <- qtm(World %>% simplify_shape(fact = 0.2), title = "Simplify 0.2")
tm4 <- qtm(World %>% simplify_shape(fact = 0.5), title = "Simplify 0.5")
tmap_arrange(tm1, tm2, tm3, tm4)

# ---
# show different options for keeping smaller (sub)units
tm5 <- qtm(World %>% simplify_shape(keep.units = TRUE, keep.subunits = TRUE),
title="Keep units and subunits")
tm6 <- qtm(World %>% simplify_shape(keep.units = TRUE, keep.subunits = FALSE),
title="Keep units, ignore small subunits")
tm7 <- qtm(World %>% simplify_shape(keep.units = FALSE),
title="Ignore small units and subunits")
tmap_arrange(tm5, tm6, tm7)
```


---

# Tool functions (colors)

## get_brewer_pal

```{r}
get_brewer_pal("Blues")
get_brewer_pal("Blues", contrast = c(.4, .8))
get_brewer_pal("Blues", contrast = c(0, 1))
get_brewer_pal("Blues", n = 15, contrast = c(0, 1))
```

```{r}
get_brewer_pal("RdYlGn")
get_brewer_pal("RdYlGn", n=11)
get_brewer_pal("RdYlGn", n=11, contrast = c(0, .4))
get_brewer_pal("RdYlGn", n=11, contrast = c(.4, 1))
```

```{r}
get_brewer_pal("Set2", n = 12)
get_brewer_pal("Set2", n = 12, stretch = FALSE)
```

## map_coloring

```{r}
data(World, metro)
World$color <- map_coloring(World, palette = "Pastel2")
qtm(World, fill = "color")
```

```{r}
# map_coloring used indirectly: qtm(World, fill = "MAP_COLORS")
data(NLD_prov, NLD_muni)
tm_shape(NLD_prov) +
  tm_fill("name", legend.show = FALSE) +
  tm_shape(NLD_muni) +
  tm_polygons("MAP_COLORS", palette = "Greys", alpha = .25) +
  tm_shape(NLD_prov) +
  tm_borders(lwd = 2) +
  tm_text("name", shadow = TRUE) +
  tm_format(format = "NLD", # Format specified for maps of the Netherlands(為荷蘭地圖指定的格式)
            title = "Dutch provinces and\nmunicipalities", 
            bg.color = "white")
```

## palette_explorer

```{r}
# 會在瀏覽器開啟一個新的視窗
# palette_explorer()
```

---

# Input and output functions

## geocode_OSM

```{r}
geocode_OSM("India")
geocode_OSM("CBS Weg 1, Heerlen")
geocode_OSM("CBS Weg 1, Heerlen", projection = 28992)
```

```{r}
data(metro)

# sample 5 cities from the metro dataset
five_cities <- metro[sample(length(metro), 5), ]
```

```{r}
# obtain geocode locations from their long names
five_cities_geocode <- geocode_OSM(five_cities$name_long, as.sf = TRUE)
five_cities_geocode
```

```{r}
# 有差異是因為地名相同。

# ---
# change to interactive mode
current.mode <- tmap_mode("view")
# plot metro coordinates in red and geocode coordinates in blue
# zoom in to see the differences
tm_shape(five_cities) +
  tm_dots(col = "blue", size = 0.1) +
  tm_shape(five_cities_geocode) +
  tm_dots(col = "red")
# restore current mode
tmap_mode(current.mode)
```

## read_GPX

沒有例子

## read_osm

Choropleth with OSM background

```{r eval = F}
# load Netherlands shape
data(NLD_muni)

# read OSM raster data
# 這個函式目前不能用惹~因為他沒有遵守一些 OpenStreetMap的規則
# [Issue with generating maps in RStudio : / Questions and Answers / OpenStreetMap Forum](https://forum.openstreetmap.org/viewtopic.php?id=68079)
osm_NLD <- read_osm(NLD_muni, ext = 1.1)
```

```{r}
# plot with regular tmap functions

# tm_shape(osm_NLD) +
# tm_rgb() +
tm_shape(NLD_muni) +
  tm_polygons("population", 
              # convert2density =TRUE, 
              style = "kmeans", 
              alpha = .7, 
              palette = "Purples")
```

A close look at the building of Statistics Netherlands in Heerlen
近距離觀察海爾倫荷蘭統計局的大樓

下面的例子已經失效，底圖都只會抓取基本的osm。

```{r}
# create a bounding box around the CBS (Statistics Netherlands) building
CBS_bb <- bb("CBS Weg 11, Heerlen", width = .003, height = .002)

# read Microsoft Bing satellite and OpenCycleMap OSM layers
CBS_osm1 <- read_osm(CBS_bb, type = "bing")
CBS_osm2 <- read_osm(CBS_bb, type = "opencyclemap")

# plot OSM raster data
qtm(CBS_osm1)
qtm(CBS_osm2)
```

## rev_geocode_OSM

```{r}
data(metro)

# sample five cities from metro dataset
set.seed(1234)
five_cities <- metro[sample(length(metro), 5), ]
```

```{r eval = F}
# obtain reverse geocode address information
# 程式碼失敗~ 因為沒有照 osm 的規則。

# ---
addresses <- rev_geocode_OSM(five_cities, zoom = 6)
five_cities <- sf::st_sf(data.frame(five_cities, addresses))
```

```{r}
# change to interactive mode
current.mode <- tmap_mode("view")
tm_shape(five_cities) +
  tm_markers(text="name")

# restore current mode
tmap_mode(current.mode)
```


---

# Others useful finctions

## calc_densities

```{r}
data(NLD_muni)
NLD_muni_pop_per_km2 <- calc_densities(NLD_muni,
                                       var = c("pop_men", "pop_women"), 
                                       target = "km km")
NLD_muni_pop_per_km2 %>% head
```

```{r}
NLD_muni <- sf::st_sf(data.frame(NLD_muni, NLD_muni_pop_per_km2))
tm_shape(NLD_muni) +
  tm_polygons(c("pop_men_km.2", "pop_women_km.2"),
              title = expression("Population per " * km^2), style = "quantile") +
  tm_facets(free.scales = FALSE) +
  tm_layout(panel.show = TRUE, panel.labels = c("Men", "Women"))
```

## get_neighbours

沒有例子

---

# END

